---
title: "Assignment 2: Coding Basics"
author: "Student Name"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
install.packages("tidyverse")
install.packages("dplyr")

seq(1,100, by = 4)
test_numbers <- seq(1,100, by = 4)

## I am using the sequence command to create a sequence of numbers from 1 to 100
## then I saved the sequence to "test_numbers"

mean(test_numbers)
median(test_numbers)

mean <- mean(test_numbers)
median <- median(test_numbers)

## here I just used the mean and median functions built into R to calculate the value for the sequence
## then I saved the mean and median values to make the next part of code easier

mean > median

## the above code will only print a TRUE value is the left value (the mean) is greater than the right value (the median)
## since it printed a FALSE value, we know the mean is not greater than the median
```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}

student <- c("Sarah", "Max", "George", "Emily") ## character vector
grade <- c(92,56,87,42) ## numeric vector
final <- c(TRUE, TRUE, FALSE, TRUE) ## logical vector

class2023 <- data.frame(student, grade, final)

library(dplyr)
class2023 <- class2023 %>% rename("Student Name" = "student", "Class Grade" = "grade", "Passed Final" = "final")
print(class2023)

```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: This is different than a matrix because it contains different classes of data, while a matrix could only contain one class. 

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r}

ifelse(grade > 50, "True", "False")

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: 'ifelse' worked because we were testing to see if a value was above '50' - if it was, it was then assigned a "TRUE" but if it was not above a '50' then it was assigned a "FALSE". So the function had to look for an 'if' and then an 'else' if that 'if' was not true. 
